---
title: "Examples DWB model"
author: "Camila Garcia and David Zamora"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Examples DWB model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# DWBmodelUN examples in a tropical basin

## DWBCalculator function

`DWBCalculator` is the function responsible for incorporating the DWB model calculations into the **DWBmodelUN** package. In this connection, the following example corresponds to an ungauged modeling exercise developed for the Sogamoso River basin (SRB) by (Duque, 2018).

As a first step, the example starts with loading datasets generated for SRB that are required by DWB model and subsequently, the coordinates system of the input data are verified. The last step is to run `DWBCalculator` function in order to obtain a monthly streamflow simulated.

```{r echo=TRUE, results='hide'}
# library(DWBmodelUN)
library(terra)
# library(raster)
# Load P and PET databases
# data(P_sogamoso, PET_sogamoso)
# # Verify that the coordinates of the databases match
# Coord_comparison(P_sogamoso, PET_sogamoso)
# # Load geographic info of GRU and parameters per cell
# data(GRU, param)
# GRU <- terra::rast(GRU)
# # Construction of parameter maps from values by GRU
# GRU.maps <- buildGRUmaps(GRU, param)
# alpha1_v <- GRU.maps$alpha1
# alpha2_v <- GRU.maps$alpha2
# smax_v <- GRU.maps$smax
# d_v <- GRU.maps$d
# 
# # Establish the initial modeling conditions
# init <- init_state(GRU.maps$smaxR)
# g_v <- init$In_ground
# s_v <- init$In_storage
# rm(init)
# 
# # Load general characteristics of modeling
# setup_data <- readSetup(Read = TRUE)
# Dates <- seq(as.Date( gsub('[^0-9.]','',colnames(P_sogamoso)[3]), format = "%Y.%m.%d"), 
#              as.Date(gsub('[^0-9.]','',tail(colnames(P_sogamoso),1)) , format = "%Y.%m.%d"), by = "month")
# Start.sim <- which(Dates == setup_data[8,1]); End.sim <- which(Dates == setup_data[10,1])
# # Sim.Period: the 1st two columns of the P and PET are the coordinates of the cells
# Sim.Period <- c(Start.sim:End.sim)+2  
# 
# # Run DWB model
# DWB.sogamoso <- DWBCalculator(P_sogamoso[ ,Sim.Period], 
#                               PET_sogamoso[ ,Sim.Period],
#                               g_v, s_v, alpha1_v, alpha2_v, smax_v, d_v)
```



## References
Duque, Nicolás. 2018. “Estimación de Campos de Precipitación En Cuencas Hidrográficas Colombianas Con Escasez de Datos, Combinando Datos Teledetectados Y de Estaciones En Tierra, Utilizando Funciones de Kernel.” Master’s thesis, Universidad Nacional de Colombia - Sede Bogotá. http://bdigital.unal.edu.co/71663/.

Tolson, Bryan A, and Christine A Shoemaker. 2007. “Dynamically dimensioned search algorithm for computationally efficient watershed model calibration” 43: 1–16.

